---
title: 'MSCS 264: Homework #12'
subtitle: 'Due Thurs May 10 at 11:59 PM'
output: md_document
---

You should submit a knitted pdf file on Moodle, but be sure to show all of your R code, in addition to your output, plots, and written responses.

```{r, setup, include=FALSE}
library(tidyverse) 
library(stringr)
```

##Factors, Functions, and Vectors

1. Given the output below, write out the elements of the vector `y`.

```{r}
#> sort(y)
#[1] -2 -2 -1  0  3  4  4  7
#
#> order(y)
#[1] 1 8 4 5 6 2 7 3
```

```{r}
y <- c(-2, 4, 7, -1, 0, 3, 4, -2)
sort(y)
order(y)
```

2. Explain what the following function does and demonstrate by running `foo1(x)` with a few appropriately chosen vectors `x`.

```{r}
foo1 <- function(x) {
  diff <- x[-1] - x[1:(length(x) - 1)]
  length(which(diff < 0))
}
```

foo1 finds the differnce between successive values of `x` and counts how many of those differences are negative (e.g. drop from the 5th value to the 6th)

3. Write a function using `if`, `else if`, and `else` to print out "good morning" if it's before 12 PM, "good afternoon" if it's between 12 PM and 5 PM, and "good evening" if it's after 5 PM.  Your function should work if you input a time like: `greet(time = "2018-05-03 17:38:01 CDT")` or if you input the current time with `Sys.time()`.

```{r}
hours <- parse_number(str_sub(Sys.time(), 12, 13))
greet <- function(time = Sys.time()) {
  hour <- parse_number(str_sub(time, 12, 13))
  if (hour < 12) {
    print("good morning")
  } else if (hour < 17) {
    print("good afternoon")
  } else {
    print("good evening")
  }
}
greet()
greet(time = "2018-05-03 17:38:01 CDT")
```

4. Write a for loop to compute the mean of every column in `mtcars`.

```{r}
output <- vector("double", ncol(mtcars))  # 1. output
for (i in seq_along(mtcars)) {            # 2. sequence
  output[[i]] <- mean(mtcars[[i]])        # 3. body
}
output
```

##Airbnb Data

5. `airbnbData.csv` contains data scraped on 2022 airbnb units in the Boston area (65 variables).  The code below shows a function which returns a set of summary statistics for a single variable.  Note a couple of strange elements (e.g. `enquo` and `!!var`) that were needed to get around the way tibbles and dplyr handle variable names (for way more detail than you probably want on this issue, see https://cran.r-project.org/web/packages/dplyr/vignettes/programming.html). 

a) Create a new function called `eda2stats` which produces the same set of summary statistics by group.  Show your function works by running `eda2stats(airbnb.df, Price, S_PropType)`.

b) Run `eda2stats` again after forming a new Property Type variable with only 4 levels: Apartment, House, Bed & Breakfast, and Other.

```{r}
airbnb.df <- read.csv("datasets/airbnbData.csv")
airbnb.df <- as.tibble(airbnb.df) 

# First remove 5 units with prices over 1000
airbnb.df <- airbnb.df %>%
  filter(Price < 1000)

# Find summary stats on a single variable
eda1stats <- function(df, var) {
  var <- enquo(var)
  df %>%
    summarise(mean = mean(!!var, na.rm = TRUE),
              median = median(!!var, na.rm = TRUE),
              sd = sd(!!var, na.rm = TRUE),
              iqr = IQR(!!var, na.rm = TRUE),
              n = n())
}
eda1stats(airbnb.df, Price)

# Find summary stats by a grouping variable
eda2stats <- function(df, var, groups) {
  var <- enquo(var)
  groups <- enquo(groups)
  df %>%
    group_by(!!groups) %>%
    summarise(mean = mean(!!var, na.rm = TRUE),
              median = median(!!var, na.rm = TRUE),
              sd = sd(!!var, na.rm = TRUE),
              iqr = IQR(!!var, na.rm = TRUE),
              n = n())
}
eda2stats(airbnb.df, Price, S_PropType)

# Create fewer levels of the grouping variable
airbnb.df <- airbnb.df %>%
  mutate(S_PropType2 = fct_lump(S_PropType, n = 3))
eda2stats(airbnb.df, Price, S_PropType2)
```

6. Use `stringr` functions to extract at least one feature from `AboutListing` and create a plot to determine if that feature is related to `Price`.  A couple of examples are illustrated below.

```{r warning=FALSE}
# Analyze features of AboutListing description
foo1 <- airbnb.df$AboutListing[5]
str_length(foo1)
str_count(foo1, "!")
str_detect(foo1, "[Ww]alk")
str_length(airbnb.df$AboutListing)[1:50]   # missings seems to have length 7

airbnb.df <- airbnb.df %>%
  mutate(NumExclaim = str_count(AboutListing, "!"),
         PctExclaim = ifelse(str_length(AboutListing) == 7, NA, 
           str_count(AboutListing, "!") / str_length(AboutListing)),
         Walk = str_detect(AboutListing, "[Ww]alk"))
airbnb.df %>% group_by(Walk) %>% tally()
airbnb.df %>% group_by(NumExclaim) %>% tally()
with(airbnb.df, summary(PctExclaim))

ggplot(airbnb.df, aes(x = Walk, y = Price)) +
  geom_boxplot() + coord_flip()
ggplot(airbnb.df, aes(x = NumExclaim, y = Price)) +
  geom_point(size = 0.25) + 
  geom_smooth(method = "lm", colour = "red")
ggplot(airbnb.df, aes(x = PctExclaim, y = Price)) +
  geom_point(size = 0.25) + 
  geom_smooth(method = "loess", colour = "red")
```
